### 2023-03-04 Preliminary Analysis/Data Processing.qmd
Added quarto data preprocessing document for preliminary analysis (still blank)

---
title: "Data Processing"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

# About this Document

This is the standing document used for preliminary data analysis for the Listening Effort in Dysarthria project. This document contains all code for processing data from pilot data collection.

```{r}

```

You can add options to executable code like this

```{r}
#| echo: false

```

The `echo: false` option disables the printing of code (only output is displayed).