### 2023-03-04 Preliminary Analysis/Data Processing.qmd
quarto document edit

---
title: "Data Processing"
format: html
editor: visual
---

# About this Document

This is the standing document used for preliminary data analysis for the Listening Effort in Dysarthria project. This document contains all code for processing data from pilot data collection.

```{r}

```

You can add options to executable code like this

```{r}
#| echo: false

```

The `echo: false` option disables the printing of code (only output is displayed).