### 2023-07-27 Peliminary Data Analysis and Vizualization.Rmd
Beginning preliminary data visualizations

---
title: "Preliminary Data Analysis and Visualization"
author: "Micah E. Hirsch"
date: "2023-07-27"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Document Purpose

The purpose of this document is to conduct preliminary analyses on pilot data for the author's dissertation project. This preliminary Analysis is based on data collected from 3 participants: LE01, LE02, and LE04. Something went wrong with LE05's transcriptions, so their data is not included in these visualizations.

```{r, include = F, warning = F, message = F}

# Load packages

library(rio)
library(tidyverse)

# import dataset

cleaned_df <- rio::import("Cleaned Data/cleaned_data.csv")

```

# Visualizations

Below are a few plots showing pupil dilation over time.

## Overall Pupil Dilation Tracking

```{r, warning = F, message = F}

cleaned_df %>%
  # For this graph, we are only looking at when the listener was accurate
  dplyr::filter(rep_acc == "accurate") %>%
  dplyr::group_by(timebins, speaker) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      color = speaker,
      fill = speaker) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,3000)) +
  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Speaker", color = "Speaker") +
  theme_bw() + 
  theme(legend.position = "bottom")

```

## Comparing Accurate and Inaccurate Responses

```{r, warning = F, message = F}

cleaned_df %>%
  dplyr::filter(subject != "LE05") %>% #There is an issue with LE05 at the moment
  dplyr::group_by(timebins, speaker, rep_acc) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      color = rep_acc,
      fill = rep_acc) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,3000)) +
  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Repetition Accuracy", color = "Repetition Accuracy") +
  theme_bw() + 
  facet_wrap("speaker") +
  theme(legend.position = "bottom")

```
## By Participant

```{r}

cleaned_df %>%
  dplyr::filter(rep_acc == "accurate") %>%
  dplyr::group_by(timebins, speaker, subject) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      fill = speaker,
      color = speaker) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,3000)) +
  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Speaker", color = "Speaker") +
  theme_bw() + 
  facet_wrap("subject") +
  theme(legend.position = "bottom")

```