### 2024-02-12 MSC 2024 Analysis/msc_analysis.Rmd
pupil dilation analysis

---
title: "Listening Effort MSC Analysis"
author: "Micah E. Hirsch"
date: "2024-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This analysis was conducted using R version 4.3.1

```{r, echo = F, warning = F, message = F}

# Load Needed Packages
library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
# Need remotes package to install gazer and saccades
library(remotes) # install.packages("remotes")
library(gazer) # remotes::install_github("dmirman/gazer")
library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
library(zoo) # install.packages("zoo")
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(ggridges) # install.packages("ggridges")
library(lme4) # install.packages("lme4")
library(lmerTest) # install.packages("lmerTest")
library(optimx) # install.packages("optimx")

```

# Listener Demographics

```{r, warning = F, message = F}

listener_demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv") |>
  dplyr::mutate(gender = factor(gender, levels = c("Man", "Woman", "Nonbinary", "Questioning", "Prefer not to answer")),
                ethnicity = factor(ethnicity, levels = c("Hispanic/Latino(a/e)", "Not Hispanic/Latino(a/e)", "Prefer not to answer")),
                race = factor(race, levels = c("white/Caucasian", "Black/African American", "Asian/Asian American",
                                               "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
                                               "Biracial or Multiracial", "Prefer not to answer", "Race not listed")),
                native_lang = factor(native_lang, c("American English", "Not American English")))

demo_table <- listener_demo |>
  dplyr::select(age, gender, ethnicity, race, native_lang) |>
  tbl_summary(type = list(age ~ "continuous",
                          gender ~ "categorical",
                          ethnicity ~ "categorical",
                          race ~ "categorical",
                          native_lang ~ "categorical"),
              statistic = list(all_continuous() ~ "{mean} ({sd}, {min}-{max})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(everything() ~ c(2)),
              label = list(age ~ "Age",
                           gender ~ "Gender",
                           ethnicity ~ "Ethnicity",
                           race ~ "Race",
                           native_lang ~ "Native Language")) |>
  as_gt()

demo_table

demo_table |>
  gt::gtsave("Tables/demo_tabe.html")

```

```{r, echo = F, warning = F, message = F}

rm(demo_table, listener_demo)

```

# Speaker Intelligibility and Phrase Accuracy

## Mean Speaker Intelligibility

```{r, warning = F, message = F}

data <- rio::import("Cleaned Data/cleaned_data_normed.csv") |>
  # removing trials with missing rep accuracy
  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) |>
  dplyr::filter(!is.na(rep_acc)) |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                trial_c = trial - 6,
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")))

speaker_acc <- data |>
  dplyr::select(subject, trial, speaker, correct_words, target_number, rep_acc) |>
  dplyr::distinct() |>
  dplyr::mutate(intel = (correct_words/target_number)*100) |>
  dplyr::select(-c(subject, trial, correct_words, target_number)) |>
  tbl_summary(by = speaker,
              type = list(intel ~ "continuous",
                          rep_acc ~ "categorical"),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(everything() ~ c(2)),
              label = list(intel ~ "Intelligibility",
                           rep_acc ~ "Repetition Accuracy")) |>
  as_gt()

speaker_acc

speaker_acc |>
  gt::gtsave("Tables/speaker_intel.html")

```

```{r, echo = F, warning = F, message = F}

rm(speaker_acc)

```

# Perceived Listening Effort

## Descriptives

```{r}

perceived_effort_df <- data |>
  dplyr::select(subject, trial, speaker, effort_rating, rep_acc, code, trial_c) |>
  dplyr::distinct() |>
  # recoding missing effort ratings (negative effort ratings indicate missing data)
  dplyr::mutate(effort_rating = ifelse(effort_rating < 0, NA, effort_rating))

perceived_effort_df |>
  dplyr::select(speaker, effort_rating, rep_acc) |>
  tbl_strata(
    strata = rep_acc,
    ~.x |>
      tbl_summary(
        by = speaker,
        type = list(effort_rating ~ "continuous"),
        statistic = list(all_continuous() ~ c("{mean} ({sd})")),
        missing = "no",
        digits = all_continuous() ~ 2,
        label = list(effort_rating ~ "Perceived Listening Effort")
      ))

```

## Visualization

```{r}

# Creating Custom Color Palette for Figures
library(NatParksPalettes) # install.packages("NatParksPalettes")
my_pal <- natparks.pals("Glacier", type = "discrete")

# Getting Sum stats for perceived effort

per_effort_des <- perceived_effort_df |>
  dplyr::group_by(speaker, rep_acc) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n()))

# Plotting the mean perceived effort ratings for the ALS and control speaker
per_effort_des |>
  ggplot() +
   aes(x = rep_acc,
       y = per_effort,
       group = speaker,
       color = speaker,
       fill = speaker) +
  geom_bar(stat = "identity", alpha = 0.6, position = position_dodge()) +
  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
  labs(x = "Accuracy", y = "Perceived Listening Effort Rating") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  scale_color_manual(values = c(my_pal[1], my_pal[3])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[3]))

# Saving the ggplot

ggsave("Plots/Perceived_effort.png", plot = last_plot(), height = 5, width = 5, units = "in")

```

## Perceived Listening Effort Models

### Fully Unconditional Model

```{r}

m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = perceived_effort_df)

sjPlot::tab_model(m0, pred.labels = "Intercept",
                  dv.labels = "Perceived Listening Effort")

performance::icc(m0)

```
### Model 1

```{r}

m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m1, pred.labels = "Intercept",
                  dv.labels = "Perceived Listening Effort")

anova(m0, m1)

```

### Model 2

```{r}

m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), 
              data = perceived_effort_df)

sjPlot::tab_model(m2, pred.labels = c("Intercept", "Trial Order"),
                  dv.labels = "Perceived Listening Effort")

anova(m1, m2)

```
### Model 3

```{r}

m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m3, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

anova(m2, m3)

```
### Model 4

```{r}

m4 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

anova(m3, m4)

```
### Model 5

```{r}

m5 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m5, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

anova(m4, m5)

```
### Model 6

```{r}

m6 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + speaker*trial_c + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m6, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Trial Order"),
                  dv.labels = "Perceived Listening Effort")

anova(m5, m6)

```

### Model 7

```{r}

m7 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
                speaker*rep_acc + speaker*trial_c + rep_acc*trial_c + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Trial Order",
                                      "Phrase Repetition Accuracy [Inaccurate] * Trial Order"),
                  dv.labels = "Perceived Listening Effort")

anova(m6, m7)

```

### Model 8

```{r}

m8 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
                speaker*rep_acc + speaker*trial_c + rep_acc*trial_c + speaker*rep_acc*trial_c + (trial_c|subject) + (trial_c|code), 
           data = perceived_effort_df)

sjPlot::tab_model(m8, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Trial Order",
                                      "Phrase Repetition Accuracy [Inaccurate] * Trial Order", 
                                      "Phrase Repetition Accuracy * Speaker * Trial Order"),
                  dv.labels = "Perceived Listening Effort",
                  file = "Tables/perceived_effort_model.html")

anova(m7, m8)

```

```{r, echo = F, message = F, warning = F}

rm(m0, m1, m2, m3, m4, m5, m6, m7, m8, per_effort_des)

```

# Pupil Dilation

## Preliminary Visualizations

Needing to spot any participants with odd trackings. Based on the visualizations below, LE13 and LE23 appear to have odd trackings, so they will be dropped from the current analyses.

```{r, warning = F, message = F}

data |>
  ggplot() +
  aes(x = time_norm,
      y = normed_pupil) +
  geom_smooth() +
  coord_cartesian(xlim = c(0,4908), ylim = c(-100, 200)) +
  facet_wrap("subject")

```

## Descriptives

### Getting Time Landmarks

```{r, message = F, warning = F}

data <- data |>
  dplyr::filter(subject != "LE13" & subject != "LE23")

data |>
  # Selecting relevant parameters
  dplyr::select(subject, trial, speaker, time_norm, code) |>
  # Filtering out times before phase onset
  dplyr::filter(time_norm >= 0) |>
  dplyr::group_by(speaker, code) |>
  dplyr::summarize(length = max(time_norm) - min(time_norm)) |>
  dplyr::ungroup() |>
  dplyr::summarize(av_length = mean(length),
                   av_phrase = av_length - 3000,
                   av_end_roi = av_length - 2000) %>%
  gt::gt() |>
  gt::cols_label(av_length = "Average Trial Length (ms)",
                 av_phrase = "Average Phrase Length (ms)",
                 av_end_roi = "Average Endpoint for ROI (ms)")

```

## Visualizations

```{r, warning = F, message = F}

data |>
  ggplot() +
  aes(x = time_norm,
      y = normed_pupil,
      group = speaker,
      color = speaker) +
  geom_smooth() +
  geom_vline(xintercept = 0, size = 0.35) +
  geom_vline(xintercept = 1910, linetype = 2) +
  annotate("rect", xmin = 500, xmax = 2910, ymin = -50, ymax = 200, alpha = 0.2) +
  coord_cartesian(xlim = c(0,4910), ylim = c(-100, 200)) +
  labs(x = "Normalized Time (ms)", y = "Pupil Dilation (arbitrary units)",
       caption = "Grey Box: Analysis Region of Interest \nSolid Line: Phrase Onset \nDotted Line: Phrase Offset") +
  facet_wrap("rep_acc") + 
  theme_bw() +
  scale_color_manual(values = c(my_pal[1], my_pal[3])) +
  theme(aspect.ratio = 1,
        plot.caption = element_text(hjust = 0))

ggsave("Plots/Pupil_dilation.png", plot = last_plot(), width = 10, height = 5, units = "in")

```

## Pupil Dilation Models

### Unconditional Growth Curve Model

```{r}

# Getting region of interest for analysis
data_roi <- data |>
  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
  dplyr::filter(time_norm >= 500) |>
  dplyr::group_by(subject, trial) |>
  # filtering out timebins 1000 ms after phrase offset
  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
  ungroup()

# Adding linear, quadratic, and cubic orthogonal polynomials to the df
data_roi <- gazer::code_poly(data_roi, predictor = "time_norm", poly.order = 3, orthogonal = T, draw.poly = F)

# Unconditional Growth Model
m0 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 +
                (poly1 + poly2 + poly3|subject) + (poly1 + poly2 + poly3|code), 
              data = data_roi)

sjPlot::tab_model(m0, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term"),
                  dv.labels = "Pupil Dilation")

performance::icc(m0)

```
### Model 1

```{r}

# Running into convergence issues, so removed the random effects for phrase (i.e. code)
m1 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m1, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]"),
                  dv.labels = "Pupil Dilation")

anova(m0, m1)

```

### Model 2

```{r}

m2 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m2, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Pupil Dilation")

anova(m1, m2)

```
### Model 3

```{r}

m3 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m3, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term"),
                  dv.labels = "Pupil Dilation")

anova(m2, m3)

```

### Model 4

```{r}

m4 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m4, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term"),
                  dv.labels = "Pupil Dilation")

anova(m3, m4)

```
### Model 5

```{r}

m5 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m5, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term"),
                  dv.labels = "Pupil Dilation")

anova(m4, m5)

```
### Model 6

```{r}

m6 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m6, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Pupil Dilation")

anova(m5, m6)

```
### Model 7

```{r}

m7 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 +
                (poly1 + poly2 + poly3|subject), 
              data = data_roi)

sjPlot::tab_model(m7, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term"),
                  dv.labels = "Pupil Dilation")

anova(m6, m7)

```

### Model 8

```{r}

m8 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 + rep_acc*poly2 +
                (poly1 + poly2|subject), 
              data = data_roi)

sjPlot::tab_model(m8, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term"),
                  dv.labels = "Pupil Dilation")

anova(m7, m8)

```

### Model 9

```{r}

m9 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
                (poly1 + poly2|subject), 
              data = data_roi)

sjPlot::tab_model(m9, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term"),
                  dv.labels = "Pupil Dilation")

anova(m8, m9)

```
### Model 10

```{r}

m10 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
                speaker*rep_acc*poly1 +
                (poly1 + poly2|subject), 
              data = data_roi)

sjPlot::tab_model(m10, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
                                      "Speaker * Phrase Repetition Accuracy * Linear Term"),
                  dv.labels = "Pupil Dilation")

anova(m9, m10)

```
### Model 11

```{r}

m11 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
                speaker*rep_acc*poly1 + speaker*rep_acc*poly2 +
                (poly1 + poly2|subject), 
              data = data_roi)

sjPlot::tab_model(m11, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
                                      "Speaker * Phrase Repetition Accuracy * Linear Term",
                                      "Speaker * Phrase Repetition Accuracy * Quadratic Term"),
                  dv.labels = "Pupil Dilation")

anova(m10, m11)

```
### Model 12

```{r}

m12 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
                speaker*rep_acc*poly1 + speaker*rep_acc*poly2 + speaker*rep_acc*poly3 +
                (poly1 + poly2|subject), 
              data = data_roi)

sjPlot::tab_model(m12, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
                                      "Speaker * Phrase Repetition Accuracy * Linear Term",
                                      "Speaker * Phrase Repetition Accuracy * Quadratic Term",
                                      "Speaker * Phrase Repetition Accuracy * Cubic Term"),
                  dv.labels = "Pupil Dilation")

anova(m11, m12)

```

### Final Model

```{r}

final_model <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
                speaker*rep_acc*poly1 + speaker*rep_acc*poly2 +
                (poly1 + poly2|subject), 
              data = data_roi)

sjPlot::tab_model(final_model, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
                                      "Speaker * Phrase Repetition Accuracy * Linear Term",
                                      "Speaker * Phrase Repetition Accuracy * Quadratic Term"),
                  dv.labels = "Pupil Dilation", 
                  file = "Tables/pupil_dilation_model.html")

```