### 2024-05-17 Manuscript Analysis/adaptation_analysis.Rmd
added rep acc analysis

---
title: 'Exploratory Analyses: Adaptation'
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document contains the code and results for an exploratory analysis done with the data collected from this study. Specifically, this document reports findings about change in listening effort (measured by perceived listening effort ratings and pupil dilation) across the study trials.

```{r, echo = F, warning = F, message = F}

library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
library(mgcv) # install.packages("mgcv")
library(itsadug) # install.packages("itsadug")
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")

```

# Proportion of Correct Responses per Trial

```{r}

# Loading in Phrase Repetition Accuracy Data
phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::select(subject, trial, targetphrase, initial_response, target_number, correct_words_initial, rep_acc) |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate"))) |>
  dplyr::select(!c(targetphrase, initial_response))

# Loading in PLE data and merging transcription accuracy with it
ple_data <- rio::import("Cleaned Data/cleaned_ple_data.csv") |>
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                trial_c = trial - 6,
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
  dplyr::filter(!is.na(rep_acc)) |>
  dplyr::filter(effort_rating > 0)

my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")

rep_acc_df <- ple_data |>
  dplyr::select(subject, trial_c, speaker, rep_acc)

rep_acc_df |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, trial_c, rep_acc) |>
  dplyr::summarize(n = n()) |>
  dplyr::ungroup() |>
  tidyr::pivot_wider(names_from = "rep_acc",
                     values_from = "n") |>
  dplyr::mutate(accurate = ifelse(is.na(accurate), 0, accurate),
                inaccurate = ifelse(is.na(inaccurate), 0, inaccurate),
                prop_accurate = accurate/(accurate + inaccurate)) |>
  ggplot() +
  aes(x = trial_c,
      y = prop_accurate,
      group = speaker,
      color = speaker) +
  geom_line() +
  labs(y = "Proportion of Accurate Responses",
       x = "Trial Number") +
  facet_wrap("speaker", ncol = 1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(my_pal[1], my_pal[5]))

ggsave("Figures/acc_response_plot.png", plot = last_plot())

```

```{r}

rep_acc_df <- rep_acc_df |>
  dplyr::mutate(accurate = ifelse(rep_acc == "accurate", 1, 0))

rep_acc_mod <- glmmTMB(accurate ~ trial_c*speaker + (trial_c|subject), family = "binomial", data = rep_acc_df)

summary(rep_acc_mod)

sjPlot::tab_model(rep_acc_mod, transform = NULL, show.est = TRUE, show.se = TRUE,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
                                  "Trial Order * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort",
                  file = "Tables/rep_acc_model.html")

```

# Perceived Listening Effort

## PLE Models

### Model 5

```{r}

m5_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m5_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

```

### Model 6

```{r}

m6_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
                    speaker*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m6_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m5_ple, m6_ple)

```

### Model 7

```{r}

m7_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
                    speaker*trial_c + rep_acc*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m7_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m5_ple, m7_ple)

```

### Model 8

```{r}

m8_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
                    speaker*trial_c + rep_acc*trial_c +  speaker*rep_acc*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m8_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m5_ple, m8_ple)

```

```{r, echo = F, message = F, warning = F}

sjPlot::tab_model(m5_ple, m6_ple, m7_ple, m8_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker * Repetition Accuracy"),
                  dv.labels = c("Model 5", "Model 6", "Model 7", "Model 8"),
                 file = "Tables/ple_adaptation_models.html")

```

## Plot

```{r, warning = F, message = F}

ple_data |>
  dplyr::mutate(trial_c = trial_c + 1) |>
  dplyr::group_by(speaker, rep_acc, trial_c) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n())) |>
  ggplot() +
   aes(x = trial_c,
       y = per_effort,
       group = speaker,
       color = speaker) +
  geom_line() +
  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0) +
  labs(x = "Trial", y = "Perceived Listening Effort Rating") +
  facet_wrap("rep_acc", ncol = 1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]))

ggsave("Figures/ple_trial_plot.png", plot = last_plot())

```

```{r}

rm(m5_ple, m6_ple, m7_ple, m8_ple, rep_acc_df, rep_acc_mod)

```

# Pupil Dilation

```{r}

pupil_df <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv") |>
  # Merging phrase accuracy information with df
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  # Filter out trials with missing phrase repetition data |>
  dplyr::filter(!is.na(rep_acc))

pupil_roi <- pupil_df |>
  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
  dplyr::filter(time_norm >= 500) |>
  dplyr::group_by(subject, trial) |>
  # filtering out timebins 1000 ms after phrase offset
  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
  ungroup() |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
                phrase = as.factor(code),
                listener = as.factor(subject),
                condition = case_when(speaker == "Control" & rep_acc == "accurate" ~ "Control Accurate",
                                      speaker == "Control" & rep_acc == "inaccurate" ~ "Control Inaccurate",
                                      speaker == "ALS" & rep_acc == "accurate" ~ "ALS Accurate",
                                      TRUE ~ "ALS Inaccurate"),
                condition = factor(condition, 
                                   levels = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate")),
                trial_c = trial - 6) |>
  dplyr::group_by(listener, speaker) |>
  dplyr::mutate(num_points = n()) |>
  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
  ungroup()

adapt_pupil <- bam(normed_pupil ~ condition +
                  s(time_norm, by = condition, k = 20) +
                  te(time_norm, trial_c, by = condition) +
                  # Random Smooths for listener, phrase, trial
                  s(time_norm, listener, bs = 'fs', m =1) +
                  s(time_norm, phrase, bs = 'fs', m = 1),
                discrete = T,
                family = "scat",
                AR.start = start_event,
                rho = .92,
                data = pupil_roi)

summary(adapt_pupil)

plot_smooth(adapt_pupil,
            view = "time_norm",
            plot_all = "condition",
            rug = F, se = 1, rm.ranef = T)

```

```{r}

gam.check(adapt_pupil)

acf_resid(adapt_pupil)

```