### 2024-06-17 Manuscript Analysis/exploratory_figures.Rmd
Created pupil dilation and intelligibility heatmap

---
title: "Exploratory Figures"
author: "Micah E. Hirsch"
date: '`r Sys.Date()`'
output:  
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this document is to conduct and create a few exploratory analyses and figures. These analyses are unplanned and not part of the main study, but were mainly created to include in M. Hirsch's dissertation defense.

```{r}

library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
library(mgcv) # install.packages("mgcv")
library(itsadug) # install.packages("itsadug")
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(viridis) # install.packages("viridis")

```

# Listening Effort and Intelligibility

## Pupil Dilation

```{r}

phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::select(subject, trial, targetphrase, initial_response, target_number, correct_words_initial, rep_acc) |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
                intel = (correct_words_initial/target_number) * 100) |>
  dplyr::select(!c(targetphrase, initial_response))

pupil_df <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv") |>
  # Merging phrase accuracy information with df
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  # Filter out trials with missing phrase repetition data 
  dplyr::filter(!is.na(rep_acc)) |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
                phrase = as.factor(code),
                listener = as.factor(subject),
                trial_c = trial - 6) |>
  dplyr::group_by(listener, speaker) |>
  dplyr::mutate(num_points = n()) |>
  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
  ungroup()

```

```{r}

intel_pupil <- bam(normed_pupil ~ speaker +
                  s(time_norm, by = speaker, k = 10) +
                  te(time_norm, intel, k = 10) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m =1),
                discrete = T,
                family = "scat",
                AR.start = start_event,
                rho = .9341941,
                data = pupil_df)

# summary.gam(intel_pupil)

gam.check(intel_pupil)

acf_resid(intel_pupil)

```

```{r}

df_pred <- expand_grid(
  time_norm = seq(from=min(pupil_df$time_norm), 
              to=max(pupil_df$time_norm), 
              length.out = 100),
  intel = seq(from=min(pupil_df$intel), 
              to=max(pupil_df$intel), 
              length.out = 100),
  speaker = c("Control", "ALS"),
  listener = c("LE01", "LE02", "LE04", "LE05", "LE06", "LE07", "LE08", "LE09", "LE10", "LE11",
               "LE12", "LE13", "LE15", "LE16", "LE18", "LE19", "LE20", "LE21", "LE22", "LE23",
               "LE24", "LE25", "LE26", "LE27", "LE28", "LE29", "LE30", "LE31", "LE32", "LE33", 
               "LE34", "LE35", "LE36", "LE38"))

df_pred <- predict(intel_pupil, newdata = df_pred, 
                      se.fit = TRUE) |> 
  as_tibble() |>
  cbind(df_pred)

df_pred |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS"),
                                   labels = c("Neurotypical", "ALS"))) |>
  dplyr::filter(time_norm >= 0 & time_norm <= 5000) |>
  ggplot() +
  aes(x = time_norm,
      y = intel,
      group = speaker) +
  geom_tile(aes(fill = fit)) +
  geom_contour(aes(z = fit), colour = "white") +
  facet_wrap("speaker") +
  theme_classic() +
  labs(fill = "Pupil Dilation Estimate",
       x = "Normed Time (ms)",
       y = "Intelligibility (% Words Correct)") +
  scale_fill_viridis(option = "D") +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))

ggsave("Figures/intel_heatmap.png", plot = last_plot())

```

## Perceived Listening Effort (PLE) Ratings

```{r}

```

# Correlations between Peak Pupil Dilaiton and PLE Ratings

```{r}

```